The Efficient Handling of Guards in the Design of RPython's Tracing JIT

### 2012-08-13 talk/vmil2012/paper.tex
improvements to guards in the frontend section

Since tracing linearizes control flow by following one concrete execution,
the full control flow of a program is not observed.
The possible points of deviation from the trace are guard operations
that check whether the same assumptions observed while tracing
still hold during execution.
Similarly, in the case of dynamic languages guards can also encode type
assumptions.
In later executions of the trace the guards can fail.
If that happens, execution needs to continue in the interpreter.
This means it is necessary to attach enough information to a guard
to reconstruct the interpreter state when that guard fails.
This information is called the \emph{resume data}.

@@

Guards interact with optimizations in various ways.
Using many classical compiler optimizations the JIT tries to remove as many
operations, and therefore guards, as possible.
In particular guards can be removed by subexpression elimination.
If the same guard is encountered a second time in the trace,
the second one can be removed.
This also works if a later guard is weaker
and hence implied by a earlier guard.

@@

Figure~\ref{fig:trace-log} shows the optimized version of the trace in
Figure~\ref{fig:fig:unopt-trace}. Allocation removal has removed the
\lstinline{new} operation and other operations handling the instance. The
operations handle unboxed numbers now.

>>>>

Since tracing linearizes control flow by following one concrete execution,
the full control flow of a program is not observed.
The possible points of deviation from the trace are denoted by guard operations
that check whether the same assumptions observed while tracing
still hold during execution.
Similarly, in the case of dynamic languages guards can also encode type
assumptions.
In later executions of the trace the guards can fail.
If that happens, execution needs to continue in the interpreter.
This means it is necessary to attach enough information to a guard
to reconstruct the interpreter state when that guard fails.
This information is called the \emph{resume data}.

@@

Guards interact with optimizations in various ways.
Using many classical compiler optimizations the JIT tries to remove as many
operations, and therefore guards, as possible.
In particular guards can be removed by subexpression elimination.
If the same guard is encountered a second time in the trace,
the second one can be removed.
This also works if a later guard is weaker
and hence implied by an earlier guard.

@@

Figure~\ref{fig:trace-log} shows the optimized version of the trace in
Figure~\ref{fig:unopt-trace}. Allocation removal has removed the
\lstinline{new} operation and other operations handling the instance. The
operations handle unboxed numbers now.
